---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r set up, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# metasim

The goal of metasim is to simulate meta-analysis data. 

I found I was rewriting the same types of analyses. I got to thinking how to make a modular set of tools for simulating meta-anlaysis data. 

In particular, I'm interested in simulating for different values of 

- $k$, number of studies
- $\tau^2$, variation between studies
- $\varepsilon^2$, variation within a study
- numbers of trials, say 10, 100, 1000
- distributions, _and_ parameters; e.g., $\exp(\lambda = 1)$ and $\exp(\lambda = 2)$. 

## work in progress

This package is a work in progress, can't guarantee anything works as intended.  

## installation

You can install metasim from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("softloud/metasim")
```

## examples


### simulate paired sample sizes

```{r packages, message=FALSE}
# packages
library(metasim)
library(tidyverse)

# so these results are reproducible
set.seed(38) 

# I like to set.seed with my age. It makes me feel smug that I'm a middle-aged woman who codes. 

```


This is a function I have often wished I've had on hand when simulating meta-analysis data. Thing is, running, say, 1000 simulations, I want to do this for the _same_ sample sizes. So, I need to generate the sample sizes for each study and for each group (control or intervention).   

Given a specific $k$, generate a set of sample sizes. 

```{r different k}

# defaults to k = 3
sim_n() %>% knitr::kable()

sim_n(k = 3) %>% knitr::kable()

# set k to a different value

sim_n(k = 6) %>% knitr::kable()


```

Suppose we require data that mimics small cohorts, say as small as 3, and as large as 50. 

```{r small cohort}
# control upper and lower bounds
sim_n(min_n = 3, max_n = 50) %>% knitr::kable()
 
```

We expect cohorts from the same study to have roughly the same size, proportional to that size. We can control this proportion with the `prop` argument.

Suppose we wish to mimic data for which the cohorts are almost exactly the same (say becaues of classes of undergrads being split in half and accounting for dropouts).

```{r prop}
# small variation between sample sizes of studies
sim_n(k = 2, prop = 0.05, max_n = 50) %>% knitr::kable()

```

It can be useful, for more human-interpretable purposes, to display the sample sizes in wide format. 

This is also useful for calculations that convert two measures to one, say, the standardised mean difference of the control and intervention groups. 

Consider four classrooms of children, who may have one or two away for illness.

```{r wide n}
sim_n(k = 4, prop = 0.05, max_n = 30, wide = TRUE) %>%
  # from here I'm just relabelling the class variable for prettiness
  separate(study, into = c("remove", "class"), sep = "_") %>% 
  select(-remove) %>% 
  mutate(class = letters[as.numeric(class)]) %>% knitr::kable()
```


### simulation parameters

Adding a few values of $\tau$, different numbers of studies $k$, and so forth can ramp up the number of combinations of simulation parameters very quickly.  

I haven't settled on a _way_ of simulating data, and haven't found heaps in the way of guidance yet. So, this is all a bit experimental. My guiding star is packaging what I'd use right now. 

What I do always end up with is generating a dataset that summarises what I would like to iterate over in simulation. 

The `sim_df` takes user inputs for distributions, numbers of studies, between-study error $\tau$, within-study error $\varepsilon$, and the proportion $\rho$ of sample size we expect the sample sizes to different within study cohorts.

```{r sim overview}
# defaults
sim_df() 

sim_df() %>% str(1)

# only consider small values of k
sim_df(k = c(2, 5, 7)) %>% str(1)
```

For the list-column of tibbles `n`, the `sim_df` function calls `sim_n` and generates a set of sample sizes based on the value in the column `k`.

```{r sample sizes set}
demo_k <- sim_df() 

# the variable n is a list-column of tibbles
demo_k %>% pluck("n") %>% head(3)


# compare the number of rows in the dataframe in the n column with the k value
# divide by two because there are two rows for each study,
# one for each group, control and intervention
demo_k %>% pluck("n") %>% map_int(nrow) %>% head(3) / 2
demo_k %>% pluck("k") %>% head(3)

```

## simulating data

Once we have established a set of sample sizes for a given distribution, with parameters, and so forth, I usually want to generate a sample for each of those `n`. We need to adjust the value of the sampled data based on the median ratio, and whether the `n` is from a control or intervention group.

A random effect is added to account for the between study error $\tau$ and within study error $\varepsilon$. 

For meta-analysis data, we work with summmary statistics, so we drop the sample and return tabulated summary stats. 

```{r simulate a sample}
sim_stats()  %>% knitr::kable()
```

## trial

In a trial, we'd first want to simulate some data, for a given distribution, for this we use the `sim_stats` function discussed in the above section. 

With the summary statistics, we then calculate an estimate of the effect or the variance of the effect. 

1. simulate data
2. calculate summary statistics
3. **calculate estimates using summary statistics**
4. calculate effects using estimates (difference, standardised, log-ratio)[^1]
5. meta-analyse 
6. return simulation results of interest

[^1]: Ideally this would be configurable but let's hardcode it for now. 


The first two steps are taken care of by the `sim_stats` function. The third step will by necessity be bespoke. 

But the rest could be automated, assuming there are the same kinds of results. 

step | input | output |
-|-|-
calculate estimates | summary statistics as produced by `sim_n` | summary stats
calculate effects | summary stats | `effect` and `effect_se`
meta-analyse | `effect` and `effect_se` | `rma` object 
summary stats | `rma` object | some kind of `broom`ing script

```{r}
metatrial()
```

## summarising simulation results

So, now we can put together some generic summarisations. Things I always want to do. Like calculate the coverage probability, confidence interval width, and bias. Most results here are mean values across all trials, the exceptions being `cp_` variables.  

`metasim` calls `metatrial` many times and summarises the results.

```{r}
metasim()
```

## simulate over parameters

```{r}
# metasims is not working yet.
```

