---
# PLEASE SEE THE README for in depth description github.com/brentthorne/posterdown
#
#---POSTER SIZE & DEFAULT FONT---#
poster_height: "38in" # height in inches of poster
poster_width: "45in" # width in inches of poster
font_family: "palatino" # choose from typical latex fonts (example: "palatino")
font_size: "30pt" #please see github.com/brentthorne/posterdown for compatible options.

#---TITLE BOX OPTIONS---#
#ESSENTIALS
title: '\fontfamily{phv}\selectfont escaping batpigday'
author: "Charles T. Gray^1^"
affiliation: "^1^Department of Mathematics and Statistics, La Trobe University, Victoria, Australia"
#STYLE & FORMATTING
titlebox_bgcol: "008080"  #Colour of the Title Box background
titlebox_bordercol: "0b4545" #Colour of the title Box border.
titlebox_shape: "all"
titlebox_borderwidth: "1cm"
title_textcol: "ffffff" #colour of title text
author_textcol: "0b4545" # Colour of author text
affiliation_textcol: "FFFFFF" # Colour of affiliation text
title_textsize: "Huge"         # Poster title fontsize
author_textsize: "Large"       # Author list font size
affiliation_textsize: "large"  # Affiliation font size
#ADDING LOGOS
logoleft_name: 'Figures/batpig'
logoleft_width: '3in'
logoleft_xshift: '1in'
logoleft_yshift: '1in'
logoright_name: 'Figures/batpig'
logoright_width: '3in'
logoright_xshift: '-1in'
logoright_yshift: '1in'

#---POSTER BODY OPTIONS---#
body_bgcol: "ffffff" #colour of the poster main background
body_textsize: "normalsize"    # Size of the main poster body text
body_textcol: "000000" # Colour of main text in the body of poster
column_numbers: 3 # Number of columns that the poster has
column_margins: "0.5in" # Margin spacing for columns
columnline_col: "008080" #colour 
columnline_width: "0pt" #width of line between each column
#SECTION TITLE STYLING
sectitle_textcol: "ffffff" # Colour of the poster section titles
sectitle_bgcol: "0b4545" # Colour of the section title box
sectitle_bordercol: "0b4545" # Colour of the border around the section title box.
sectitle_borderwidth: "2mm" # Thicknes of the section title box border
sectitle_boxshape: "uphill" # Changes the shape of the section title box.

#---BIBLIOGRAPHY OPTIONS---#
bibliography: MyLibrary # name of the .bib file used for referencing
bibliography_spacing: 0.8 # sets the multiplier for line spacing of bibliography spacing (between 0 and 1)
bibliography_textsize: "small"  # size of the bibliography text size (handy for one too many references!)

#---OTHER---#
cite_col: "CC0000" #colour of ciation elements
url_col: "008080" #colour of url links
link_col: "008080" #colour of other links within the poster
footnote_textcol: "ffffff" # Colour of footnote text if used
output: posterdown::posterdown_pdf
---

```{r chunk opts, include=FALSE}
knitr::opts_chunk$set(
  eval = TRUE,
  message = FALSE,
  echo = TRUE,
  results = 'asis',
  fig.height = 3.5,
  fig.width = 4.5,
  out.width = "100%",
  warning = FALSE,
  fig.align = 'center',
  message = FALSE,
  dev = 'cairo_pdf'
)

```

<!--
Bellow you will find typical RMarkdown syntax which should fill in the columns for the poster from top to bottom left to right. The more information you have the more will be filled. Please let me know if you have feedback at brentthorne18@gmail.com or at github.com/brentthorne/posterdown.
ENJOY! XD
-->

# batpigday

\begin{description}
  \item[batpigday]{\emph{noun} The coding equivalent of groundhogday.}
\end{description}

# the problem

***
\begin{center}
Simulating data is a bitch. 
\end{center}
***

Debugging frequently dominates the time of students in mathematical science. 
These students know how to solve equations, and next to nothing about code. 

***

New tools are emerging daily to enable researchers to avoid these timesink pitfalls. 

These tools have lowered the programmatic barrier for researchers, but it still a learning curve. 

***

We consider a case study in meta-analysis.

\begin{description}
\item[meta-analysis] Statistical methodology for combining the results of several studies. 

\end{description}

# meta-analysis of medians

Conventional meta-analytic tools, such as `metafor::rma`, require an **effect** and a **variance** of that effect.

***

But what if the reported statistics are **median** and **interquartile** range?

***

To test our proposed estimator for the variance of the sample median, I found myself repeating tasks and checks in the algorithms. 

***

I tried to find a better way of debugging and writing simulations.

This lead to: 

1. a packaged analysis, `varameta::`*, which is built on 
2. the simulation package for meta-analysis data, `metasim::`*. 

(*in development)


\vfill\null
\columnbreak

# escaping batpigday

\begin{center}
\textbf{coding is the easiest part of coding}
\end{center}

*** 

 - Modular code, break the code into chunks. 
 - Reproducibility is more than `set.seed()`: accessibility, refactoring, integratability, versioning, 
 

***

```{r load metasim, echo=TRUE}
library(tidyverse)
library(metasim)
```

```{r output table, echo=FALSE}
library(kableExtra)

# table styling
output_table <- function(df) {
  df %>% 
    janitor::adorn_rounding(skip_first_col = FALSE) %>% 
  kable(align = "c", caption = "", booktabs = TRUE) %>% 
  kable_styling(latex_options = c("striped","HOLD_position"),full_width = TRUE, font_size = 25)
}
 
```


***
Generate sample sizes for $k$ studies. 

```{r n, echo=TRUE, eval=TRUE}
# simulate 2 studies where most have at most 25
sim_n(k = 2, min_n = 10, max_n = 25) %>% output_table()

```

```{r df}
# generate simulation dataframe
sim_df() %>% head(2) %>% select(-n) %>% output_table()
```

Each **row** of this dataframe represents a set of **simulation** parameters. 

***

Each simulation runs a **trial** function. 

```{r trial}
metatrial() %>% output_table()
```
        
Each **simulation** runs the trial function |trials| times.

```{r test}
metasim() %>%
  pluck("results") %>%
  select(-coverage_count) %>%
  output_table()

```

For all **simulations**, run `metasim` over each row of the dataframe. 

```{r}
metasims() %>% output_table()

```



\vfill\null
\columnbreak


# the maths


# references

This poster was created with `posterdown::`.

<!--- Here you can set the size of the citation text as well as remove the "References" section if you choose not to have one for some reason :) -->
\small\printbibliography


