---
title: "captions"
author: "Charles T. Gray"
date: "20/05/2019"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = TRUE,
  echo = TRUE)

```

```{r packages, message=FALSE}

# packages
library(tidyverse)
library(metasim)
library(kableExtra)
library(dplyr)
library(latex2exp)

# set winners
conflicted::conflict_prefer("filter", "dplyr")
conflicted::conflict_prefer("group_rows", "kableExtra")

```


## simulate

```{r run simulation}
sim <- metasims(single_study = TRUE)

```


```{r parmaeter extractor}
sim_parameter <- function(sim, arg){

  sim %>% 
    purrr::pluck("arguments") %>% 
    dplyr::filter(argument == arg) %>% 
    purrr::pluck("value")
}


```


```{r parameters}
# parameters 
(unequal_effect_ratio <- sim %>% sim_parameter("unequal_effect_ratio") %>% 
   as.numeric())

trials <- sim %>% sim_parameter("trials") %>% as.numeric()


```


## Caption drafts


## only for meta-analysis

```{r meta-analysis parameters}

(k <- sim %>% sim_parameter("k") %>% parse(text = .) %>% eval())

(tau_sq_true <- sim %>% sim_parameter("tau_sq_true") %>% parse(text = .) %>% eval())

```

# distributions caption

b) Data for the control and intervention arms were sampled from distributions: the normal distribution, with parameters 2 and 0.3, the exponential distribution, with parameters 2, the Pareto distribution, with parameters 3 and 3, the Pareto distribution, with parameters 2 and 1, the Pareto distribution, with parameters 0.5 and 1, the log-normal distribution, with parameters 1 and 0.3. 


# plot caption

a) These simulation results are summarised by the proportion, coverage, of `r trials` confidence intervals that contain the true measure $\nu$, or log-ratio $\log(\rho)$ of control $\nu_C$ and intervention $\nu_I$ measures, with $\rho := \nu_I/\nu_C$. 

##  Caption for the simulation

The case where the control median $\nu_C$ is equal to the intervention median $\nu_I$ is considered, $\rho= 1$, as is an unequal ratio, $\rho =$ `r unequal_effect_ratio`. 

Data for the control and intervention arms sampled from distributions in accompanying Distributions plot. 

Summary statistics were calculated from the samples, for the estimator of the variance $\mathcal V(m)$ of the sample median. 

### meta-analysis only

For each of the `r trials` trials, $K$ studies' control and intervention arms were simulated.  A meta-analysis was performed via `metafor::rma` with the Knapp-Hartung test for... todo:. The confidence interval that informs the coverage was extracted from the model results. 

Simulations, `r trials` trials, were performed for each of $K =$ `r k` studies. 

The cases where there where there is no variation between the studies, $\tau^2 = 0$, and some variation introduced by independent factors other than the measure of interest for a given study,  $\tau^2 =$  `r tau_sq_true[-1]` are provided.

```{r}
# todo tau2
```


### single-study simulation
  
For each of the `r trials` trials, a confidence interval, $\theta \pm \Phi^{-1}(0.975)\mathcal \sqrt(V(\theta))$, was calculated, where $\theta$ estimates the true measure, or log-ratio of control and intervention measures.      


# experiments

## results as is
```{r results = "asis", echo = FALSE}
sim %>% plot()

sim %>% caption()


```

## fig.cap=caption(sim) + results = asis

```{r fig.cap=caption(sim), results='asis'}

sim %>% plot()

```


## fig.cap=kable(caption(sim))

```{r fig.cap=kable(caption(sim)), fig.height=0.5}

sim %>% plot()

```

### output text

<div align="center">


```{r echo=TRUE}
sim %>% caption() %>%  kable()
```



</div>
